Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-02-05 cluster_paper/ms.tex
update paper

\newcommand{\nautorot}{1{,}276\ } % drivers.plot_full_kinematics_X_rotation; 2021.01.29
\newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes

@@

\title{
  Cluster Difference Imaging Photometric Survey. III.
  %The Coeval Core \& Halo of NGC\,2516
  %The Halo of NGC\,2516 Extends To At Least 250\,pc
  An Open Cluster With a 250\,pc Halo
  %NGC\,2516  250\,pc
}

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these ``halos'' could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a 250\,pc halo reported
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
  tidal radius of the cluster is {\bf $\sim$XX\,pc}.  Using photometry
  from Gaia, rotation periods from TESS, and lithium measurements from
  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
  coeval with the core.  Two in three kinematically selected halo
  members 25-250\,pc from the cluster center show rotation periods
  consistent with a gyrochronological age of 120\,Myr.  Stars in the
  inner 25\,pc show higher period detection rates than the
  halo (73\% vs.\ 66\%), but both vastly exceed our field star
  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
  are also anti-correlated at fixed stellar mass, as is the case with
  comparable clusters.  This work expands the set of confirmed
  NGC\,2516 members by a factor of $\approx$2, and quantifies the
  degree to which Gaia-based analyses can identify dispersed stellar
  populations against a background of field star contaminants.  We
  highlight implications for spectroscopic survey targeting, open
  cluster dispersal, and planet searches around young stars.
\end{abstract}
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023

@@

\section{A 250\,pc Halo around a Core?}
\label{sec:gaia}
%
% What is the core? What is the halo?
%

@@

Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
on this plot are the \nautorot light curves for which the peak Lomb Scargle
periodogram period was below 15 days, the LS power exceeded 0.08, and
at most one equal-brightness companion was within the aperture.  These
selection criteria are entirely heuristic.  
But splitting the sample into the ``core'' and the ``halo'', they
suggest that the halo is real.

@@

The results are shown in Figure~\ref{fig:physical_x_rotn}.
The halo extends to about 250\,pc in physical space, and out to at
least 10 of km/s in tangential velocity space.
\citet{meingast_2021} by comparison required a physically motivated cut
in tangential velocity space of 1.5 km per sec.
This probably improves purity, but clearly doesn't get everthing for
completeness.

@@

		\subfloat{
			\includegraphics[width=0.75\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium equivalent widths in NGC\,2516,}
  plotted against extinction-corrected color ({\it top}), and the
  combination of rotation period and extinction-corrected color ({\it
  bottom}).
  \label{fig:lithium}
  }
\end{figure*}

>>>>

\newcommand{\nautorot}{1{,}144\ } % % depends on which cuts are adopted. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt

\newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes

@@

\title{
  Cluster Difference Imaging Photometric Survey. III.
  %The Coeval Core \& Halo of NGC\,2516
  %The Halo of NGC\,2516 Extends To At Least 300\,pc
  An Open Cluster With a 300\,pc Halo
  %NGC\,2516  300\,pc
}

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these ``halos'' could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a 300\,pc halo reported
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
  tidal radius of the cluster is {\bf $\sim$XX\,pc}.  Using photometry
  from Gaia, rotation periods from TESS, and lithium measurements from
  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
  coeval with the core.  Two in three kinematically selected halo
  members 25-250\,pc from the cluster center show rotation periods
  consistent with a gyrochronological age of 120\,Myr.  Stars in the
  inner 25\,pc show higher period detection rates than the
  halo (73\% vs.\ 66\%), but both vastly exceed our field star
  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
  are also anti-correlated at fixed stellar mass, as is the case with
  comparable clusters.  This work expands the set of confirmed
  NGC\,2516 members by a factor of $\approx$2, and quantifies the
  degree to which Gaia-based analyses can identify dispersed stellar
  populations against a background of field star contaminants.  We
  highlight implications for spectroscopic survey targeting, open
  cluster dispersal, and planet searches around young stars.
\end{abstract}
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023

@@

\section{A 300\,pc Halo around a Core?}
\label{sec:gaia}
%
% What is the core? What is the halo?
%

@@

Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
on this plot are the \nautorot light curves for which the peak Lomb Scargle
periodogram period was below 15 days, the LS power exceeded 0.08, and
for which no equal-brightness companions were within the aperture.  These
selection criteria are entirely heuristic.  
But splitting the sample into the ``core'' and the ``halo'', they
suggest that the halo is real.

@@

The results are shown in Figure~\ref{fig:physical_x_rotn}.
The halo extends to separations of about 200\,pc in physical space
from the cluster core.
This corresponds to a length of $\approx$300-350\,pc, depending on
which members of the halo are chosen as the ``tips'' on either
end.

In tangential velocity space, the fraction of stars with rotation
period detections remains high out to roughly 10\,km/s.
\citet{meingast_2021} by comparison required a physically motivated
cut in tangential velocity space of 1.5 km per sec.  This probably
improves purity, but clearly doesn't get everthing for completeness.

@@

		\subfloat{
			\includegraphics[width=0.75\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium equivalent widths in NGC\,2516,}
  plotted against extinction-corrected color ({\it top}), and the
  combination of rotation period and extinction-corrected color ({\it
  bottom}).
  {\bf TODO: Do this against the autorot sample; Add the GALAH EWs}
  \label{fig:lithium}
  }
\end{figure*}
